---
title: "Data analysis"
bibliography: reference.bib
format:
  html:
    theme: default
    code-fold: true
    code-summary: "Show the code"
---


![Source: [@us_maps]](us_map.jpeg)

## Question
This analysis aims to gather some information about different states in the US. I used to think the capital is most likely to be the largest city in the state, however it seems most of the time it is not true, and after reading the article[@states_admission], I am think whether is has some relation with the history. Therefore, I conducted this analysis try to explore the relationship between total area, history of each state, and its population.

## Dataset Intro
The dataset is from tidytuesday[@tidytuesday]. The dictionary and dataset can be found on this [website](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-08-01/readme.md).
After installing the tidytuesdayR, people also can get access to this dataset.

## Audiance  
This data analysis is a demonstration of the visualization skills. It should be further developed using more advanced techniques. Based on that, the current audience is the students in this course.

::: {#about-block}

```{r}
library(tidytuesdayR)
library(tidyverse)
library(dplyr)
library(ggplot2)
```


# Import Data

```{r}
states= tidytuesdayR::tt_load('2023-08-01')
states_general=states[["states"]]
state_name_etymology=states[["state_name_etymology"]]
```
:::

## Data Wrangling
```{r}
data_final=states_general%>%
  full_join(state_name_etymology, by="state")%>%
  distinct(state, .keep_all=TRUE)%>%
  mutate(matched_cities=if_else(capital_city==largest_city,1,0))%>%
  mutate(population_2020=population_2020/1000000)%>%
  mutate(total_area_km2=total_area_km2/1000)%>%
  select(-total_area_mi2,-land_area_mi2, -water_area_mi2)

data_final%>%
  head(5)
```

::: {.callout-tip}
Some state have more than one name and meanings, so we have over 50 row in our original dataset. After use distinct function, there should be only 50.
:::

```{r}
state_land=data_final%>%
  select(state, land_area_km2)%>%
  rename(area=land_area_km2)%>%
  mutate(area_type="Land")
state_water=data_final%>%
  select(state, water_area_km2)%>%
  rename(area=water_area_km2)%>%
  mutate(area_type="Water")
water_land=state_water%>%
  bind_rows(state_land)%>%
  print()
```

::: {.callout-note}
This smaller dataset contains the 50 states and their land and water area, we will use this to explore the common fact about 50 states.
:::

## Part 1
```{r,fig.width=8}
data_final%>%
  ggplot(aes(x=admission, y=population_2020, colour=factor(matched_cities)))+
  geom_point()+
  geom_text(aes(label=postal_abbreviation),size=2,vjust = -0.5)+
  labs(title = "The date when the state admitted to the union vs. the population",
       subtitle = " Although, we can't see the name of each points, there is no relationship between the date the state admitted to the union. Also whether the capital is the largest city has no correlation with the population", x="The date when the state admitted to the union.", y="The population of state in 2020(in million)", 
       color="whether the largest cities is the captial", caption="Chuyi Sun")

```

```{r}
#| column: margin
knitr::kable(
  data_final%>%
    select(state, admission)%>%
    arrange(admission)%>%
    head(5)
)
```

## Part 2

```{r, fig.width=8}
data_final%>%
  ggplot(aes(x=total_area_km2, y=population_2020))+
  geom_smooth(se=F)+  
  labs(title = "The relationship between land and population",
       subtitle = "The relationship between population and area of the state is not a strict linear relationship", x="total area in thounds km2", y="total poplation(in million)", caption = "Chuyi Sun")+
  theme(axis.text = element_text(size =12), axis.title = element_text(size = 14))
  
```

```{r}
#| column: margin
knitr::kable(
  data_final%>%
    select(state, total_area_km2)%>%
    arrange(desc(total_area_km2))%>%
    head(5)
)
```

```{r}
#| column: margin
knitr::kable(
  data_final%>%
    select(state, population_2020)%>%
    arrange(desc(population_2020))%>%
    head(5)
)
```


## Part 3
```{r, fig.width=9, fig.height=18}

water_land%>%
  ggplot(aes(x=area_type, y=area, fill=area_type))+
  geom_bar(stat="identity")+
  facet_wrap(~state,scales = "fixed", nrow=10)+
  labs(title = "The water area and land area of each state",
       subtitle = "Overall, all the state have more land area than water area. Alaska has the largest land area and water area", x="the type of area", y="area in thounds km2", caption = "Chuyi Sun")


```

## Part 4 

##### Summary 

Before the analysis, I assumed there might be a relationship between the total area, history of each state, and population. However, from the results, we can't establish any relationship between these three aspects. The additional results we can conclude is US has more land area than the water area, which is  consistent with its latitude and longitude. Lastly, if the state has the more area that doesn't mean that the population is large, eg Alaska[@alaska] has a large area, but the climate is not good for living.

![Source: [@alaska]](Alaska.jpeg)

## Part 5

###### Functions used

dplyr

- mutate
- distinct
- select
- if_else
- filter
- inner_join
- bind_rows
- rename

ggplot2 

- geom_point
- geom_bar
- geom_smooth

